---
title: "TitalGPUReproduction"
author: "Samyoga Bhattarai"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
The paper uses survival analysis methods (Kaplan-Meier + Cox). So, installing and loading the required libraries.
```{r}
#installing required packages
#install.packages(c("tidyverse", "lubridate", "Survival", "survminer", "forestmodel"))

#loading the packages
library(tidyverse)
library(lubridate)
library(survival)
library(survminer)
library(forestmodel)
```


Loading the dataset:
The csv file (gpu_summary_loc.csv) is a pre-processed GPU lifetime data. So, directly using this for the survival analysis technique
```{r}
data <- read.csv("./data/gc_summary_loc.csv")

str(data)
summary(data)
```

Defining the required variables for survival analysis:
Here, the two columns 'years' and 'dead' from the dataset are crucial variables as years shows the lifetime of each GPU and dead is the event where 1 shows GPU failure and 0 shows GPU did not fail so censored. The event can be Dual Bit Error (DBE) or Off The Bus (OTB).
```{r}
data$time <- data$years
data$event <- data$dead
```

```{r}
#cleaning the data to avoid missing values
data <- data %>% filter(!is.na(time), time > 0)
```
