---
title: "TitalGPUReproduction"
author: "Samyoga Bhattarai"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries
The paper uses survival analysis methods (Kaplan-Meier + Cox). The paper also does Time Between Failure analysis. So, installing and loading the required libraries.
```{r}
#installing required packages
#install.packages(c("tidyverse", "lubridate", "Survival", "survminer", "forestmodel"))
#install.packages(c("scales", "patchwork"))

#loading the packages
library(tidyverse)
library(lubridate)
library(survival)
library(survminer)
library(forestmodel)
library(scales)
library(patchwork)
```


## 1. Load the dataset

Two datasets are used: `gc_summary_loc.csv` and `gc_full_csv`.

The csv file (gc_summary_loc.csv) is a pre-processed GPU lifetime data. So, directly using this for the survival 
analysis technique
```{r}
# Summary dataset: one row per GPU serial number
data <- read.csv("./data/gc_summary_loc.csv")

str(data)
summary(data)
```

```{r}
# Full dataset: one row per GPU-location lifetime segment
gc_full <- read.csv("./data/gc_full.csv")
str(gc_full)
```

Defining the required variables for survival analysis:
Here, the two columns 'years' and 'dead' from the dataset are crucial variables as years shows the lifetime of each GPU and dead is the event where 1 shows GPU failure and 0 shows GPU did not fail so censored. The event can be Dual Bit Error (DBE) or Off The Bus (OTB).
```{r}
#survival time and event variables
data$time <- data$years
data$event <- data$dead
```

```{r}
#cleaning the data to avoid missing values
data <- data %>% filter(!is.na(time), time > 0)
```

Converting Predictors to categorical variables using factor for cox model:
```{r}
data$col <- as.factor(data$col)
data$row  <- as.factor(data$row)
data$cage <- as.factor(data$cage)
data$slot <- as.factor(data$slot)
data$node <- as.factor(data$node)

#adding batch variable and converting to categorical for considering old and new batch. Old batch - inserted before 2016 and new batch - inserted in 2016 or later.
data$batch <- as.factor(data$batch)

cat("Total GPUs in analysis:", nrow(data), "\n")
cat("Events (failures):", sum(data$event), "\n")
cat("Censored:", sum(data$event == 0), "\n")
cat("\nBatch breakdown:\n")
print(table(data$batch))
cat("\nFailures by batch:\n")
print(table(data$batch, data$event))
```

Creating survival object:
```{r}
#double bit error object for survival analysis
surv_dbe_obj <- Surv(time = data$time, event = data$dead_dbe)
#OTB error object for survival analysis
surv_otb_obj <- Surv(time = data$time, event = data$dead_otb)
#combined DBE and OTB for survival analysis
surv_obj <- Surv(time = data$time, event = data$event)
```


Kaplan-Meier Estimation:
```{r}
kaplanmeier_fit <- survfit(surv_obj ~ cage, data = data)

plot_kaplanmeier <- ggsurvplot(
  kaplanmeier_fit,
  data = data,
  conf.int = TRUE,
  risk.table = TRUE,
  ggtheme = theme_minimal(),
  risk.table.height = 0.30
)

print(plot_kaplanmeier)
```

Cox proportional hazard model:
```{r}
cox_model <- coxph(Surv(time, event) ~ col + row + cage + slot + node, data = data
)

summary(cox_model)
```

Interpreting hazard ratios:
```{r}
exp(coef(cox_model))
```
```{r}
forest_model(cox_model)
```

Schoenfeld residual test
```{r}
ph_test <- cox.zph(cox_model)
ph_test
plot(ph_test)
```

```{r}
summary(cox_model)
```

Adjusted survival curves by cage
```{r}
newdata_adj <- expand.grid(
  cage = levels(data$cage),
  col  = levels(data$col)[1],
  row  = levels(data$row)[1],
  slot = levels(data$slot)[1],
  node = levels(data$node)[1]
)

fit_adj <- survfit(cox_model, newdata = newdata_adj)

ggsurvplot(
  fit_adj,
  data = newdata_adj,
  conf.int = TRUE,
  ggtheme = theme_minimal(),
  legend.title = "Cage"
)
```

```{r}
ggsurvplot(
  kaplanmeier_fit,
  data = data,
  fun = "cloglog",
  ggtheme = theme_minimal()
)
```

